---
title: "Homework 1"
author: "Zo Kupelian"
date: "2026-09-18"
output: pdf_document
---

```{r}
library(dplyr)
```


#1.Loading tlc dataset

```{r}
install.packages("haven")
library(haven)

tlc <- read_sas("tlc.sas7bdat")
```

#2. Select just the placebo group (trt = P)

```{r}
just_placebo <- tlc %>%
  subset(trt == "P")
```

#3. Find mean and variance of Pb (lead) at y0, y1, y4, and y6

```{r}
#I could write a function and lapply through this, but this is simple and I'm lazy

y0_mean <- mean(just_placebo$y0)
y0_var <- var(just_placebo$y0)
y0_mean
y0_var

y1_mean <- mean(just_placebo$y1)
y1_var <- var(just_placebo$y1)
y1_mean
y1_var

y4_mean <- mean(just_placebo$y4)
y4_var <- var(just_placebo$y4)
y4_mean
y4_var

y6_mean <- mean(just_placebo$y6)
y6_var <- var(just_placebo$y6)
y6_mean
y6_var
```

#4. Find covariance matrix of y0, y1, y4, and y6

```{r}

```

#5. Confidence interval for the difference of mu_1 - mu_0 from estimated means from step 3
##a. Assuming covariance is 0

```{r}

```

##b. Using covariance from step 4

```{r}

```

#6. Two-sample t test comparing y1 and y0

```{r}

```

#7. Paired t test of the change from week 0 to week 1 calculated in step 7

```{r}

```

#8. Compare CI 

A: R's means by hand covariance = 0
B: R's means by hand covariance from step 4
C: SAS's means by hand covariance = 0
D: SAS's means by hand covariance from step 4
E: From R's two-sample t test comparing y1 with y0
F: From R's paired t test of change from week 0 to week 1
G: From SAS's two-sample t test comparing y1 with y0
H: From SAS's paired t test of change from week 0 to week 1

