---
title: "Homework 2"
author: "Zo Kupelian"
date: "2026-09-25"
output: pdf_document
---

```{r}
library(dplyr)
library(tidyverse)
library(nlme)
```

#1.Loading tlc dataset

```{r}
#install.packages("haven")
library(haven)
library(nlme)

tlc <- read_sas("tlc.sas7bdat")
```

#2. Select just the placebo group (trt = P)

```{r}
just_placebo_long <- tlc %>%
  subset(trt == "P") %>%
  pivot_longer(starts_with("y"), names_to = "time", values_to = "y")
just_placebo_long
```

#3. Fit incorrect general linear model

```{r}
wrong_model <- lm(y ~ relevel(as.factor(time), 4), data = just_placebo_long)
summary(wrong_model)
anova(wrong_model)
```
##R Results
F statistic: 2.2088
DOF: 196
P-value: 0.088

Fail to reject the null hypothesis; there's insufficient evidence to say that
the means differ across the different time points.

##SAS Results

F statistic: 2.21
DOF: 196
P-value: 0.0884

Fail to reject the null hypothesis; there's insufficient evidence to say that
the means differ across the different time points.

#4. Fit generalized linear model

```{r}
right_model <- gls(y ~ relevel(as.factor(time), 4),
                   data = just_placebo_long,
                   correlation = corSymm(form = ~ 1|id),
                   weights = varIdent(form = ~1|time),
                   method = "ML"
                   )
summary(right_model)
anova(right_model)
```
##R Results

F statistic: 10.2321
DOF: 196
P-value: <0.0001

Reject the null hypothesis with the P value under 0.05, so there is evidence for
the means differing across time points.

##SAS Results

F statistic: 10.23
DOF: 196
P-value: <0.001


#5. Question about what a piece of output from SAS and nlme in R is about.

For the nlme summary, I understood most of the output, but I wasn't entirely 
sure what the "Formula: ~1 | time" output meant. 

From SAS, the entire "Iterations History" table was something I didn't recognize.
Iteration 1 or 0 didn't seem to match up to anything in the data or the tests
we were running.

#6. Revert whole dataset to include trt = A as well as P

```{r}
tlc_longer <- tlc %>%
  pivot_longer(starts_with("y"), names_to = "time", values_to = "y")
```

#7. Fit model with both groups and follow the process listed in Canvas.

```{r}
#same thing as question 3, just with tlc_longer
model1 <- gls(y ~ relevel(as.factor(time), 4) * relevel(as.factor(trt), 2), 
              data = tlc_longer,
              correlation = corSymm(form = ~ 1 | id),
              weights = varIdent(form = ~ 1 | time),
              method = "ML"
              )

summary(model1)
anova(model1)
```

We're seeing significance, which means we reject the null hypothesis; there IS
a difference over time. This trend over time is NOT parallel.

```{r}
# 1. Look at the parameter estimates from model1
summary(model1)

# This shows the coefficients for each time point and treatment group.
# The interaction coefficients show how the treatment effect differs at each time point.

# 2. Get predicted values for each group at each time point

# Create a data frame with all combinations of time and treatment
newdata <- expand.grid(
    time = c("y0", "y1", "y4", "y6"),
    trt = c("P", "A")
)

# Get predictions
predictions <- predict(model1, newdata = newdata)

# Combine predictions with the newdata
results <- cbind(newdata, predicted_y = predictions)
print(results)

# 3. Visualize the trajectories
ggplot(results, aes(x = time, y = predicted_y, color = trt, group = trt)) +
    geom_point(size = 3) +
    geom_line() +
    labs(title = "Predicted Response Profiles by Treatment Group",
         x = "Time Point",
         y = "Mean Response",
         color = "Treatment") +
    theme_minimal()
```

The output of the SAS graph is very similar to the one here in R - the two 
groups start out nearly identical at y0, but the active group sees a massive 
dip, then a strong increase immediately after the treatment is applied. The 
placebo group, on the other hand, has a slow decrease in the average response 
rate of the group as time goes on.

The groups are significantly different, which means the treatment does have an
effect.
